Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-04-18 manuscript.tex
Ms corrections
Ms reorganization almost to submit

@@

\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}

@@

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano
et al. 2005; Thompson et al. 2012). Trophic networks (or food webs)
allow identifying properties and key species that may be crucial for
ecosystem stability, and hence important for ecosystem management and
biodiversity conservation (Thompson et al. 2012). There is a growing
interest in understanding how ecological networks respond to
environmental and anthropogenic stressors along geographical gradients
(Cirtwill et al. 2015; Bauer et al. 2022). Yet, only a few studies have
described variation in food web structure along latitudinal gradients in
marine ecosystems. The few that have come from the Global North (Wood et
al. 2015; Kortsch et al. 2019; Pecuchet et al. 2022), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. We recruited food web
experts from different marine systems of Argentina and the world (see
co-authors' list). Throughout the year 2023 we maintained regular
discussion meetings, typically held every two or three weeks. The aim of
this review is threefold: 1) describe the complexity and structure of
marine food webs along the southwest Atlantic to Antarctic axis from a
network perspective; 2) identify the ongoing environmental and
anthropogenic stressors for each marine ecosystem containing the food
webs; and 3) review proven food web changes and elaborate hypotheses on
how the identified stressors might affect food web features (e.g.~energy
flow, stability), combining information on node- and network-level
properties. Finally, we suggest which additional data and analyses are
needed to gain insights into the stressors' effects on food web
properties in the southwest Atlantic to Antarctic region.

\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha et al. 2004; Latorre et al. 2023). The referred region
extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (Matano et al. 2010; Guihou et al. 2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa and Varisco 2007; Pasotti et al.
2015; Saunders et al. 2019; Riccialdelli et al. 2020), but few studies
have considered the complexity of the ecosystem in terms of a high
resolution of species and their prey-predator interactions (but see
Jacob et al. (2011), Marina et al. (2018), Funes et al. (2022),
López-López et al. (2022), Rodriguez et al. (2022), Marina et al. (in
rev.)). Neglecting this complexity might lead to a misunderstanding of
the structure and functioning of the ecosystems, and ultimately reduce
the ability to predict ecosystem responses to perturbations (Montoya et
al. 2009).

In this review, we consider marine areas in the Southern hemisphere for
which highly-resolved empirical food webs, in terms of species diversity
and trophic interactions, have been previously described. These areas
include: (1) San Jorge Gulf (45 - 47ºS, 65 - 68ºW), (2) Beagle Channel
(\textasciitilde54ºS, 68ºW), (3) Burdwood Bank (\textasciitilde54ºS,
59ºW), (4) Scotia Sea (48 - 58ºS, 50ºW), (5) Potter Cove (62ºS, 58ºW,
Antarctica), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) (Figure
1). The geographic locations of these marine biomes span from temperate
to Antarctic, and are exposed to both environmental (sea warming,
glacial retreat) and anthropogenic (fishery, pollution) stressors.

@@

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Glembocki et al. 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (Góngora et al.
2012), as well as marine mammal and seabird populations (Yorio 2009).
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic
interactions with a connectance of 0.04. The percentage of top predators
is 16\%, 78\% of intermediate nodes and 6\% of basal nodes; 60\% of
predators are omnivorous (Table 1). The most connected nodes are: the
Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea
gregaria, squids (Illex argentinus as dominant species) and Amphipoda.
Notably, these nodes present mid-trophic positions in the food web (3.0,
2.5, 3.6 and 2, respectively) (Funes et al. 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al. 2023). Bruno et al. (2023) suggests that
locally produced suspended particulate organic matter (mostly composed
by phytoplankton) and organic matter accumulated in the sediments
(macroalgae-originated detritus) are the primary food sources for the
marine ecosystem, as opposed to allochthonous materials. Beagle Channel
food web includes 145 nodes and 1115 trophic interactions with a
connectance of 0.05 (Table 1). The food web is suggested to have a
wasp-waist structure, where the following species play a crucial role in
the dynamics of the ecosystem: Fueguian sprat Sprattus fuegensis,
longtail southern cod Patagonotothen ramsayi, black southern cod P.
tessellata, frogmouth Cottoperca trigloides, and squat lobster Grimothea
gregaria (Riccialdelli et al. 2020). Moreover, the squat lobster has
been identified as a species responsible for linking modules and
connecting the entire food web (Rodriguez et al. 2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Administración de
Parques Nacionales 2022) (Figure 1). Physical features in the plateau
are fairly stable, with salinity averaging 34 all year round and
temperature ranging between 4 and 8ºC (Acha et al. 2004). The plateau is
surrounded by steep flanks of up to 4000 m depth, protected by the
Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2;
\textasciitilde55º - 56ºS, \textasciitilde58º - 62ºW). Intense upwelling
and mixing occur in relation to the slope, entraining deep nutrient-rich
waters into the photic layer (Matano et al. 2019), and resulting in a
fairly homogeneous water column both spatially and temporally (Matano et
al. 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al. in rev.) (Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to
the West and by the island complex of the Scotia Arc to the North, East,
and South, with an approximate extension of 1.5 x 106 km2 and a depth
range of 0 - 3000 m (Murphy et al. 2006) (Figure 1). Its oceanography is
dominated by the Antarctic Circumpolar Current, which is spatially
structured by frontal systems (Whitworth 1980). The South Antarctic
Circumpolar Current Front subdivides the Scotia Sea into two
biogeographic regions: the Northern Scotia Sea is characterized by
higher and more variable temperatures, and the Southern Scotia Sea by
lower and more stable temperatures and influenced by seasonal sea ice
(Raymond 2011). The analysis of the Northern and Southern Scotia Sea
food webs shows that the former is relatively more complex than the
latter: with higher species richness (218 vs 192) and interactions
(10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As
expected from a more complex network, the path length is shorter in the
Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea
network displays a greater proportion of omnivores and a lower mean
trophic level (López-López et al. 2022) (Table 1).

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al. 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al. 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (Quartino and Boraso de Zaixso
2008). The Potter Cove food web includes 110 nodes and 649 interactions,
with an asymmetric degree distribution, and a connectance value of 0.05
(Table 1). It presents a modular structure (groups of species interact
more strongly with each other than with species belonging to other
groups), that's positively associated with stability, since perturbation
can be retained within modules (Rodriguez et al. 2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al. 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Teixidó et al. 2002). The Weddell Sea food
web exhibits a high level of network complexity, featuring the greatest
number of nodes (490) and trophic interactions (16041) among the
analyzed food webs in this review (Table 1). Its connectance (0.07) and
other structural properties are intermediate compared to the other food
webs. Recently, the interaction strengths of this food web were
estimated, revealing the presence of numerous weak and few strong
interactions, which is consistent with findings in other complex food
webs (Nilsson and McCann 2016). Notably, this asymmetric distribution of
interaction strength is likely to promote community persistence.

>>>>

\usepackage{multirow}

@@

\section{The response of trophic interaction networks to multiple
stressors along a large-scale latitudinal range in the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-along-a-large-scale-latitudinal-range-in-the-southern-hemisphere}

@@

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano
et al. 2005; Thompson et al. 2012). Trophic networks (or food webs)
allow identifying properties and key species that may be crucial for
ecosystem stability, and hence important for ecosystem management and
biodiversity conservation (Thompson et al. 2012). There is a growing
interest in understanding how ecological networks respond to
environmental and anthropogenic stressors along geographical gradients
(Cirtwill et al. 2015; Bauer et al. 2022). Yet, only a few studies have
described variation in food web structure along latitudinal gradients in
marine ecosystems. The few that have come from the Global North (Wood et
al. 2015; Kortsch et al. 2019; Pecuchet et al. 2020), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

Here we review for the first time stressor responses of marine food webs
for areas located along a Southwest Atlantic to Antarctic gradient (45 -
78ºS, Figure 1). We considered proven and expected changes in food webs
driven by global change-related stressors in selected areas along this
large-scale latitudinal range. The aim of this review is threefold: 1)
describe the complexity and structure of marine food webs along the
Southwest Atlantic to Antarctic axis from a network perspective; 2)
identify the ongoing global change-related stressors for each marine
ecosystem containing the food webs; and 3) review proven food web
changes and elaborate on a series of alternative hypotheses on how the
identified stressors might affect food web features (energy flow,
stability and topology), combining information on node- and
network-level properties. To identify the main stressors, we performed
an exploratory and configurative review instead of a systematic review
(Haddaway et al. 2015), because our objective was to formulate
evidenced-based hypotheses rather than providing an exhaustive list of
species-level stressors for each ecosystem. Finally, we suggest which
additional data and analyses are needed to gain insights into the
stressors' effects on food web properties in the Southwest Atlantic to
Antarctic region.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Assessment of the food web structure and stressors in the
Southwest Atlantic - Antarctic
region}\label{assessment-of-the-food-web-structure-and-stressors-in-the-southwest-atlantic---antarctic-region}

The Southwest and the Atlantic sectors of the Southern Ocean comprise
one of the most biologically productive regions of the world's oceans
(Acha et al. 2004; Latorre et al. 2023). The referred region extends
from San Jorge Gulf (45ºS) in the Patagonian shelf to the Weddell Sea
(78ºS) in the Southern Ocean and covers a well-connected oceanic area
(Matano et al. 2010; Guihou et al. 2020).

Throughout this region, many studies address the ecology of
predator-prey relationships but most of these studies are restricted to
a few interacting species or certain components of the food web such as
fish, crustaceans, and benthos (e.g., Vinuesa and Varisco (2007),
Padovani et al. (2012), Pasotti et al. (2015), Riccialdelli et al.
(2020), Saunders et al. (2019), Pasti et al. (2021), Cordone et al.
(2022)). Only few studies consider the complexity of trophic
interactions within highly resolved food webs containing hundreds of
species, ranging from phytoplankton, zooplankton, benthos, fish,
seabirds to whales (but see Jacob et al. (2011), Marina et al. (2018),
Funes et al. (2022), López-López et al. (2022), Rodriguez et al. (2022),
Marina et al. (in rev.)). Neglecting this complexity might lead to a
misunderstanding of the structure and functioning of the ecosystems and
ultimately reduce the ability to predict ecosystem responses to
perturbations (Montoya et al. 2009).

The Southwest Atlantic - Antarctic marine biota has been and is
currently subjected to a variety of stressors (e.g., sea warming,
glacial retreat, ice changes, acidification, species invasion,
fisheries, and contamination). Here, we consider a stressor as a global
change variable if it causes a quantifiable change, irrespective of its
direction (increase or decrease), in a biological response (Orr et al.
2020). This aligns with the framework established during the early years
of the International Geosphere-Biosphere Programme (IGBP) for
classifying driving forces. These driving forces are grouped into those
that modify the composition of the Earth's fluid envelopes (e.g., sea
warming, acidification), and those that originate as local processes but
attain global importance due to their widespread impact (e.g., species
invasion, fisheries) (Vitousek 1992; Sage 2020). Although it is
well-known that multiple stressors act in concert at any given time
(e.g., warming and fishery; acidification and contaminants) (Gutt et al.
2021), to date stressor assessments have been performed individually.
Moreover, the potential for interactive effects of two or more stressors
(synergy or antagonism) (Côté et al. 2016) is almost unknown for the
region (Rowlands et al. 2021).

In this review, we included six large marine areas in the Southern
Hemisphere for which highly-resolved empirical food webs in terms of
species diversity and trophic interactions have been previously
described: (1) San Jorge Gulf (\textasciitilde45 - 47ºS, 65 - 68ºW)
(Funes et al. 2022), (2) Beagle Channel (\textasciitilde54ºS, 68ºW)
(Rodriguez et al. 2022), (3) Burdwood Bank (\textasciitilde54ºS, 59ºW)
(Marina et al. in rev.), (4) Scotia Sea (\textasciitilde48 - 58ºS, 50ºW)
(López-López et al. 2022), (5) Potter Cove (62ºS, 58ºW, Antarctica)
(Marina et al. 2018), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica)
(Jacob et al. 2011; Marina et al. 2024) (Figure 1). The geographic
locations of these marine biomes span from temperate Atlantic waters to
Antarctic, and are exposed to global change stressors (e.g., sea
warming, glacial retreat, fishery, pollution) (Sage 2020), of which some
are locally more important, e.g., fisheries in the San Jorge Gulf or
glacial retreat in the Antarctic.

In the following subsections, we describe the complexity and structure
of the mentioned marine food webs, identify the main stressors, the
species (or trophic species), and the parameters affected. Table 1
summarizes the properties of the marine food webs (number of nodes,
trophic interactions, connectance, path length, mean trophic level, and
omnivory); Table 2 provides a list of the main stressors affecting
species inhabiting each area, considering: stressors' nature, species
and parameter affected, and locality.

@@

\subsubsection{2.1 San Jorge Gulf}\label{san-jorge-gulf}

San Jorge Gulf is the northernmost study area considered in this review
(Figure 1), located from Cabo dos Bahias (44°55'S) to Cabo tres Puntas
(47°06'S). It is a partially enclosed basin spanning approximately
34,000 km\^{}2 and \textasciitilde100 m of maximum depths. The gulf's
productivity supports large invertebrate and vertebrate fisheries
(Góngora et al. 2012), as well as marine mammal and seabird populations
(Yorio 2009). Two prominent frontal systems meet in the shallower
northern and southern ends of the gulf (with depths \textasciitilde40
m), which are also the places of highest productivity (Glembocki et al.
2015).

The San Jorge Gulf food web contains 165 nodes (trophospecies) and 1015
trophic interactions with a connectance of 0.04. The percentage of top
predators is 16\%, 78\% of the nodes are intermediate and 6\% are basal;
60\% of the consumers are omnivorous (Table 1). The most connected nodes
are the Argentine red shrimp \emph{Pleoticus muelleri}, the squat
lobster \emph{Gimothea gregaria}, squids (\emph{Illex argentinus} as
dominant species), and Amphipoda. Notably, these nodes are positioned at
mid-trophic levels in the food web (3, 2.5, 3.6, and 2, respectively)
(Funes et al. 2022).

The gulf is subject to several global change stressors (Table 2),
especially fisheries (González-Zevallos and Yorio 2006, 2011; Galván et
al. 2022). By discarding species, trawl fisheries add new trophic
interactions to the food web which results in decreasing trait
variability and stability of the system (Rincón-Díaz et al. 2021; Funes
et al. 2022). Moreover, this has changed the availability of resources
to some predators. For example, the Argentine hake \emph{Merluccius
hubbsi}, one of the main bycatch species, has become a prey item of
non-diving seabirds, like the kelp gull \emph{Larus dominicanus}
(González-Zevallos and Yorio 2006) and reef fish (Funes et al. 2019).
Although juveniles of \emph{M. hubbsi} is largely the main bycatch
species, 29 cartilaginous and 69 bony local fish species were also
registered as incidental catch between 2005 and 2014 (Bovcon et al.
2013; Ruibal Nuñez 2020). This level of impact has triggered a shift in
the functional diversity of fish assemblages in the gulf, homogenizing
their trophic function (Rincón-Díaz et al. 2021), for example in the
sites with the historical highest fishing activity, the fish assemblage
experienced a relative increase in midwater fusiform species and lost
variability in their depth range (Rincón-Díaz et al. 2021). Other
functional changes were a decrease in the maximum sizes of individual
fish, together with a drop in elasmobranch biomass and an increase in
crustacean biomass (Funes 2020). The significant increase in crustacean
biomass was mainly due to the increase in the Argentine red shrimp and
the squat lobster (Funes 2020). These species have become the most
important prey for the most abundant fish in the area, the Argentine
hake and the narrownose smooth-hound \emph{Mustelus schmitti} (Belleggia
et al. 2017; Pasti et al. 2021). However, since commercial fisheries
ceased their activities in the San Jorge Gulf in 2015 (Annex I,
Resolution CFP No 7/2018), the effects mentioned above from trawl
fisheries on the structure and function of the ecosystem may have
changed since then. Nowadays only small-scale trawling from artisanal
fishery is practiced in the gulf. Sea warming is another important
stressor in the San Jorge Gulf, because of the southward shifts of
northern fish populations to the area (Galván et al. 2022). Signs of sea
warming have been detected in the gulf and nearby regions. Anomalous
bottom temperature increments of up to 2°C have been observed in the
gulf since 2000 (Isla and Cortizo 2024), and a positive trend in sea
surface temperature of 0.3°C/0.4°C per decade over a 30-year series
(Saraceno et al. 2022; Risaro et al. 2022). Additionally, a positive
trend in sea surface temperature (\textasciitilde0.5 °C increase in the
last 20 years) has been reported for the Nuevo and San Jose Gulfs, both
located close to and northward of the San Jorge Gulf (Williams and
Nocera 2023). The region is especially prone to be affected by
climate-driven shifts in species ranges because it is situated in the
ecotone of two biogeographic provinces, the Argentine (30°S - 44°S) and
the Magellanic (43°S - 55°S) (Balech and Ehrlich 2008). In addition to
tropicalization from northern fish populations, alien species are also
documented to affect demersal assemblages of fish and macroinvertebrates
in the gulf (Galván et al. 2022). Finally, the gulf is exposed to urban
and industrial pollution (Verga et al. 2020). In particular, there is an
oil monobuoy from which oil manipulation and general oil transport along
the Patagonian coast registered several oil spills and chronic oil
discharges (García-Borboroglu et al. 2008). Other marine systems
impacted by the oil spill have shown extraordinary mortality events for
cetacean (Litz et al. 2014), sea turtles (Wallace et al. 2017), and
important decrease in marine bird populations (Irons et al. 2000) sea
otters (Garrott et al. 1993) and macroalgae (Southward and Southward
1978), with cascading effects throughout the food web (Peterson 2001).

\subsubsection{2.2 Beagle Channel}\label{beagle-channel}

Located at a more southern latitude (\textasciitilde54ºS, 68ºW), the
Beagle Channel (Figure 1) is an interoceanic passage at the southernmost
tip of South America, spanning 240 km in length and 5 km in width
(\textasciitilde1200 km\^{}2), with a depth range of 0 - 140 m. It
features complex coastlines, varying bathymetry, a prevailing
west-to-east circulation pattern, and a significant longitudinal
gradient of glacial freshwater discharge (Schloss et al. 2023). Bruno et
al. (2023) suggests that locally produced suspended particulate organic
matter (mostly composed of phytoplankton) and organic matter accumulated
in the sediments (macroalgae-originated detritus) are the primary food
sources for the marine ecosystem, as opposed to allochthonous materials.

The Beagle Channel food web includes 145 nodes and 1115 trophic
interactions with a connectance of 0.05 (Table 1). This food web is
hypothesized to have a wasp-waist structure (i.e., the structure and
dynamics of the ecosystem are regulated primarily by mid-trophic level
species), where the following species play a crucial role in the
dynamics of the ecosystem: Fueguian sprat \emph{Sprattus fuegensis},
longtail southern cod \emph{Patagonotothen ramsayi}, black southern cod
\emph{P. tessellata}, frogmouth \emph{Cottoperca trigloides}, and squat
lobster \emph{G. gregaria} (Riccialdelli et al. 2020). Moreover, the
squat lobster has been identified as a species linking food web modules
and connecting the entire food web (Rodriguez et al. 2022).

A main local stressor in the Beagle Channel is the introduction of
invasive species (Table 2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially, the Chinook salmon \emph{Oncorhynchus
tshawytscha} causes concern. The detection of this species in Tierra del
Fuego dates back to April 2006, and its population has been expanding
since then (Nardi et al. 2019). Being a top predator, the Chinook salmon
can compete for resources with several native predators in the Beagle
Channel (Correa and Gross 2008), but also prey on them. This is the case
for Notothenioids (\emph{Eleginops maclovinus}, \emph{Patagonotothen
tessellata}, \emph{P. cornucola}, \emph{P. sima}, \emph{Paranotothenia
magellanica}, \emph{Harpagifer bispinis}), the Atherinidae
\emph{Odontesthes smitti} and \emph{O. nigricans}, the Fuegian sprat,
and larvae of king crabs (\emph{Lithodes antarcticus} and
\emph{Paralomis granulosa}) (Fernández et al. 2010). Ciancio et al.
(2008) observed that Chinook salmon primarily feeds on sprats in the
Southern Patagonian Shelf Ecosystem area and display trophic levels
comparable to those of intermediate-sized fish and cephalopod predator
species, showing significant dietary overlap with the Magellanic penguin
\emph{Spheniscus magellanicus}. Another potential competitor for Chinook
salmon in the Beagle Channel, sharing a similar diet, is the native
Commerson's dolphin \emph{Cephalorhynchus commersonii} (Riccialdelli et
al. 2013). In the Beagle Channel, other stressors include contaminants
like metals, perfluorinated compounds, hydrocarbons, and microplastics
found in animal tissue (algae, mussel, fish, and guano) and sediments
(Duarte et al. 2011; Llorca et al. 2012). Carbon and carbohydrate levels
in Ushuaia Bay's (54°48′S, 68°18′W) surface sediments are similar to
hypertrophic ecosystems (i.e., greater input of nutrients than an
eutrophic situation), along with hydrocarbons and heavy metals linked to
port and industrial activities (Gil et al. 2011; Commendatore et al.
2012). Ferreira et al. (2021) showed that the black southern cod in
Ushuaia Bay may be exposed to endocrine-disrupting compounds from urban
and industrial pollution. Moreover, Pérez et al. (2020) and Ojeda et al.
(2021) found microplastics in the Chilean mussel \emph{Mytilus
chilensis} and the sea snail \emph{Nacella magellanica}, respectively.
These studies link pollutants to port and industrial activities of
Ushuaia; we therefore expect the area of the Beagle Channel closest to
the city to be the most affected by contaminants. Contaminants
potentially spread through the food web, magnifying the risk to higher
trophic level taxa. In the Beagle Channel and Atlantic coast of Tierra
del Fuego food webs mercury concentrations rose with benthivory, and
higher mercury was found in phytoplankton and the squat lobster. This
latter species connects pelagic and benthic habitats, so any disruption
to them could disrupt the food web (Rodriguez et al. 2022). Dodino et
al. (2022) found the highest mercury levels in Magellanic penguin
feathers from offshore colonies in Tierra del Fuego. Recently, Ushuaia's
kelp forests have also seen changes in composition along with a decrease
in biodiversity as a result of urban pollution (Kaminsky et al. 2023).
It is important to note that sea warming might be a less significant
stressor in the Beagle Channel compared to the other areas, as global
warming is melting glaciers and the influx of cold freshwater is
actually slightly cooling the seawater (Saraceno et al. 2022).

\subsubsection{2.3 Burdwood Bank}\label{burdwood-bank}

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II (\textasciitilde54ºS, 59ºW), meaning
the shallow submarine plateau named Burdwood Bank with a 200 m isobath
boundary, and a deep slope that reaches 4000 m in depth, respectively
(Administración de Parques Nacionales 2022) (Figure 1). Physical
features in the plateau are fairly stable, with salinity averaging 34
all year round and temperature ranging between 4 and 8ºC (Acha et al.
2004). The plateau is surrounded by steep flanks of up to 4000 m depth,
protected by the Marine Protected Area Namuncurá - Burdwood Bank II
(32,000 km\^{}2). Intense upwelling and mixing occur in relation to the
slope, entraining deep nutrient-rich waters into the photic layer
(Matano et al. 2019), and resulting in a fairly homogeneous water column
both spatially and temporally (Matano et al. 2019).

The Burdwood Bank food web comprises 379 nodes and 1788 interactions,
with a connectance of 0.01, and an asymmetric degree distribution
(i.e.~most of the species have a relatively low number of interactions
and few species concentrate most of them). Almost half of the consumers
are omnivores (48\%), and the network displays a small-world pattern
(Marina et al. in rev.) (Table 1). This network pattern implies a rapid
spread of a perturbation (e.g., contaminant, population fluctuation,
local extinction) throughout the food web due to a short distance among
species but, at the same time, a greater resistance caused by the
confinement of perturbations mainly within subnetworks as a result of a
high clustering coefficient (Bornatowski et al. 2017; Dormann et al.
2017).

The main stressors reported for the ecosystem of the Marine Protected
Areas Namuncurá within Burdwood Bank I and II are related to human
activities such as fisheries and contamination (Table 2). Several
fisheries targeting demersal fish operate in the vicinity and within the
Marine Protected Areas. The fishery of the Patagonian toothfish
\emph{Dissostichus eleginoides} has gained prominence in recent years
(Allega et al. 2020; Gorini et al. 2021). Although these are regulated
by the Argentinean government, incidental bycatches do occur, where
demersal fish of the genera \emph{Coelorinchus} and \emph{Macrourus},
seabirds and benthic invertebrates (30+ taxa) are the most common
bycatches (Gaitán and Marí 2016; Martínez et al. 2022). Noteworthy,
among the invertebrates caught, eight species are indicator taxa of
vulnerable marine ecosystems (Gaitán and Marí 2016; Schejter and Albano
2021). However, independent assessments of these bycatches suggest no
significant impact on the populations (Gaitán and Marí 2016; Martínez et
al. 2022). Besides bycatches, seabirds are being affected by fisheries
due to discards, altering their diet; the most frequently encountered
species are black-browed albatross \emph{Thalassarche melanophris},
Southern giant petrel \emph{Macronectes giganteus}, Cape petrel
\emph{Daption capense}, Southern royal albatross \emph{Diomedea
epomophora}, Northern giant petrel \emph{M. halli}, and white-chinned
petrel \emph{Procellaria aequinoctialis} (Tamini et al. 2023).
Nevertheless, there is a lack of knowledge considering their role in the
ecosystem and the potential joint effect of both target fish and bycatch
in a broader food web framework. Contaminants such as microplastics and
mercury occur in the water column of the Burdwood Bank ecosystem (Cossi
et al. 2021; Fioramonti et al. 2022; Di Mauro et al. 2022).
Microplastics are distributed along the entire water column, from
surface to deep waters (3-2450 m) (Di Mauro et al. 2022). More
importantly, microplastics were found in soft tissues of benthic
macroinvertebrates (sea stars \emph{Henricia obesa} and \emph{Odontaster
penicillatus}) and benthopelagic fish (\emph{Patagonotothen guntheri}
and \emph{P. ramsayi}), which not only incorporated the contaminant from
the environment through their filter-feeding system but could also get
it indirectly from prey organisms already containing plastics in their
tissues (Cossi et al. 2021). Notably, one of the contaminated species is
the longtail southern cod \emph{P. ramsayi}. This species is part of the
core group of species that drive the ecosystem through the suggested
wasp-waist control (Riccialdelli et al. 2020). Mercury transfer and
biomagnification are occurring at a greater pace than near coastal areas
such as the Beagle Channel (Fioramonti et al. 2022). The Fuegian sprat
\emph{S. fuegensis}, a pelagic fish with a mid-trophic level in the food
web, has the highest levels of mercury recorded (Fioramonti et al.
2022). Considering the wasp-waist control of the Fuegian sprat in the
food web (Riccialdelli et al. 2020), a rapid and widespread
contamination to the top predators is expected (Fioramonti et al. 2022).
Despite evidence of warming of surface, mid-water and bottom layers (100
m) in Burdwood Bank (Franco et al. 2020a), studies on the oceanographic
aspects of the system are lacking. Warming impacts on the species and
trophic interactions in Burdwood Bank are also unknown.

\subsubsection{2.4 Scotia Sea}\label{scotia-sea}

The Scotia Sea is a deep-sea basin (\textasciitilde48 - 58ºS, 50ºW),
delimited by the Drake Passage to the West and by the island complex of
the Scotia Arc to the North, East, and South, with an approximate
extension of 1.5 x 106 km\^{}2 and a depth range of 0 - 3000 m (Murphy
et al. 2006) (Figure 1). Its oceanography is dominated by the Antarctic
Circumpolar Current, which is spatially structured by frontal systems
(Whitworth 1980). The South Antarctic Circumpolar Current Front
subdivides the Scotia Sea into two biogeographic regions: the Northern
Scotia Sea, characterized by higher and more variable temperatures, and
the Southern Scotia Sea, characterized by lower and more stable
temperatures and influenced by seasonal sea ice (Raymond 2011).

Analysis of the Northern and Southern Scotia Sea food webs shows that
the former is relatively more complex in terms of number of species and
links than the latter: with more species richness (218 vs 192) and
interactions (10008 vs 7241) and a slightly higher connectance (0.21 vs
0.20). Mean path length is shorter in the Northern Scotia Sea food web,
whereas the Southern Scotia Sea network displays a greater proportion of
omnivores and a lower mean trophic level (López-López et al. 2022)
(Table 1).

The Scotia Sea is a vast and heterogenous oceanic region, where
especially the zone around South Georgia island represent an area of
interest, here referred to as `Northern Scotia Sea'. The majority of
studies analyzing the stressors' effects come from this area (e.g.,
Murphy et al. (2007); Whitehouse et al. (2008); Trathan (2023)) (Table
2). The main stressor of the Scotia Sea is commercial fisheries. Krill
fishery on \emph{Euphausia superba} not only impacts the targeted
species, but also the many predators that depend on it as a food source
(Hilborn et al. 2017). Yet, data currently available from monitoring of
krill and its predators remain insufficient, hence identifying the
potential fishery impacts on the ecosystem is difficult (Trathan et al.
2021). Apart from krill fishery, two other commercial fisheries operate
in the Scotia Sea, targeting Patagonian toothfish species \emph{D.
eleginoides} and \emph{Dissostichus mawsoni}. The \emph{D. eleginoides}
stock is linked to the stock at South Georgia (`Northern Scotia Sea')
(Collins et al. 2010), while the \emph{D. mawsoni} stock is linked to
the Antarctic continental shelf (`Southern Scotia Sea') (Soeffker et al.
2022). Recently, Trathan (2023) has identified several concerns
regarding aspects of these fisheries, highlighting that it is crucial to
recognize ongoing changes such as the recovery of baleen whales and
climate change. These factors increase uncertainty for fisheries
managers, necessitating direct consideration in the management of
harvested and dependent species.
